---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(dplyr)
devtools::load_all()
```

# covidReff

<!-- badges: start -->
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-green.svg)](https://www.tidyverse.org/lifecycle/#stable)
<!-- badges: end -->


The goal of `covidReff` is to simulate Covid outbreaks in a partially vaccinated
population. 

## Installation

The development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("grattan/covidReff")
```

## Usage

The `simulate_covid()` runs simulations of Covid spreading and returns a `tibble` with the results for each simulation (`runid`) on each day. Using the default settings:

```{r}
sim_results <- simulate_covid()
```

The resulting `tibble` is:

```{r}
glimpse(sim_results)
```

The key inputs of the `simulate_covid()` function -- among many -- are the reproduction value of the virus in a 'relaxed' Australian society, `R`, and the proportion of the **whole** population that are vaccinated, `vaccination_levels`. 

```{r}
sim_r8_50 <- simulate_covid(R = 8, 
                            vaccination_levels = .5)
```

The `vaccination_levels` is provided as either a single numeric for vaccination levels uniformly-distributed across age groups (as above), or you can provide a numeric vector of length `10` describing the vaccination rates for age groups `1-10`, `11-20`, `21-30`, ..., `91-100`.

```{r}
sim_r4_50 <- simulate_covid(
  R = 4, 
  vaccination_levels = c(
    "0-10"  = 0.00,
    "11-20" = 0.40,
    "21-30" = 0.60,
    "31-40" = 0.60,
    "41-50" = 0.60,
    "51-60" = 0.70,
    "61-70" = 0.90,
    "71-80" = 0.90,
    "81-90" = 0.95,
    "91+"   = 0.95)
  )
glimpse(sim_r4_50)
```
