---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

devtools::load_all()
```

# covidReff

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->


The goal of `covidReff` is to simulate Covid outbreaks in a partially vaccinated
population. 

This package is in development and should not be relied upon.

## Installation

The development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("grattan/covidReff")
```

## Usage

The `simulate_covid()` runs simulations of Covid spreading and returns a `tibble` with the results for each simulation (`runid`) on each day. Using the default settings:

```{r}
sim_results <- simulate_covid()
sim_results
```

The key inputs for the `simulate_covid()` function are `R0` and `vaccination_levels`:

```{r}
# oh no
oh_no <- simulate_covid(R = 8, vaccination_levels = .5)
oh_no
# okay!
okay <- simulate_covid(R = 2, vaccination_levels = .9)
okay
```

